HO unification from object language to meta language

### 2024-03-22 paper/main.tex
address -> addr + cleanup

\begin{figure}[H]
  \begin{tabular}{ll}
  \begin{minipage}{0.21\textwidth}
  \begin{elpicodetab}
kind fm type.
type fapp list fm -> fm.
type flam (fm -> fm) -> fm.
type fcon string -> fm.
type fuva nat -> fm.
\end{elpicodetab}
  \end{minipage}
  &
  \begin{minipage}{0.24\textwidth}
  \begin{elpicodetab}
kind tm type.
type app list tm -> tm.
type lam (tm -> tm) -> tm.
type con string -> tm.
type uva nat -> list tm -> tm.
  \end{elpicodetab}
  \end{minipage}
  \end{tabular}
  \caption{\Fo{} and \Ho language}\vspace{0.3em}
  \label{code:common-terms}
  \Description[code:common-terms]{code:common-terms}
\end{figure}

@@

\begin{elpicode}
typeabbrev memory A (list (option A)).
type set?   nat -> memory A -> A -> o.
type unset? nat -> memory A -> o.
type assign nat -> memory A -> A -> memory A -> o.
\end{elpicode}

@@

\begin{elpicode}
  typeabbrev fsubst (memory ftm).
  kind assmt type.
  type abs (tm -> assmt) -> assmt.
  type val tm -> assmt.
  typeabbrev subst (memory assmt).
\end{elpicode}

@@

@@

The auxiliary function \elpiIn{close-liks}

@@

\noindent
Note that \elpiIn{link} carries the arity (number of expected arguments) of
the variable. \marginpar{say when this is needed}

\begin{elpicode}
type solve-links links -> links -> subst -> subst -> o.
solve-links L L S S.
\end{elpicode}

Then decomp

\begin{elpicode}
type decompile links -> subst -> fsubst -> o.
decompile L S O :-
  map (_\r\r = none) S O1, % allocate empty fsubst
  (pi N X\ knil N X :- mem L (link X N _) ; N = X) =>
    decompl S L O1 O.
type knil nat -> nat -> o.

type decompl links -> subst -> fsubst -> o.
decompl S [] [].
decompl S [link _ N _|L] O P :- unset? N S X,
  decompl S L O P.
decompl S [link M N _|L] O P :- set? N S X,
  decomp-assignment S X T, assign M O (some T) O1,
  decompl S L O1 P.

type decomp-assignment subst -> assignment -> fm -> o.
decomp-assignment S (abs F) (flam G) :-
  pi x y\ decomp-tm S x y => decomp-assignment S (F x) (G y).
decomp-assignment S (val T) T1 :- decomp S T T1.

type decomp subst -> tm -> fm.
decomp _ (con C) (fcon C).
decomp S (app A) (app B) :- map (decomp S) A B.
decomp S (lam F) (flam G) :-
  pi x y\ decomp S x y => decomp S (F x) (G y).
decomp S (uva N A) R :- set? N S F,
  move F A T, decomp S T R.
decomp S (uva N A) R :- unset? N S,
  map (decomp S) A B, knil N M, napp (fapp[fuva M|B]) R.
\end{elpicode}

\marginpar{TODO link}
\marginpar{TODO nuove subst}

Now unif\marginpar{TODO: code unif}

\begin{elpicode}
~\PYG{k+kd}{type} \PYG{n+nf}{(\Uo)} \PYG{k+kt}{fm -> fm -> subst -> subst -> o}~.
(X ~\Uo~Y) S S1 :-
  fderef S X X0, fderef S Y Y0,                    ~($\customlabel{step:norm}{norm}$)~
  comp X0 X1 [] S0 [] L0,                        ~($\customlabel{step:compile}{compile}$)~
  comp Y0 Y1 S0 S1 L0 L1,
  (X1 ~\Ue~Y1) [] HS0,                              ~($\customlabel{step:unify}{unify}$)~
  solve-links L1 L2 HS0 HS1,                        ~($\customlabel{step:link}{link}$)~
  decompile L2 HS1  S1.                        ~($\customlabel{step:decompile}{decompile}$)~
\end{elpicode}

\subsection{Prolog simulation}
Allows us to express the properties.
we take all terms involved in a search (if a rule is used twice we simply
take a copy of it), we compile all of them, and then we pick the unification
prblems among these terms and step trough them.

\begin{elpicode}
type pick list A -> (pair nat nat) -> (pair A A) -> o.
pick L (pr X Y) (pr TX TY) :- nth X L TX, nth Y L TY.

type prolog-fo list fm -> list (pair nat nat) -> subst -> o.
prolog-fo Terms Problems S :-
  map (pick Terms) Problems FoProblems,
  fold4 ~(\Uo)~ FoProblems [] S.

type step-ho (pair tm tm) -> links -> links -> subst -> subst -> o.
step-ho (pr X Y) L0 L1 S0 S2 :-
  (X1 ~\Ue~Y1) S0 S1,
  solve-links L0 L1 S1 S2.

type prolog-ho list fm -> list (pair nat nat) -> subst -> o.
prolog-ho Terms Problems S :-
  fold4 comp Terms HoTerms [] L0 [] HS0,
  map (pick HoTerms) Problems HoProblems,
  fold4 step-ho HoProblems L0 L HS0 HS,
  decompile L HS S.
\end{elpicode}

\noindent
the proprty is that if a step for Fo succeds then the Ho one does, and
if Fo fails then the Ho fails ()

>>>>

\begin{figure}[H]
  \begin{tabular}{ll}
  \begin{minipage}{0.21\textwidth}
   \inputrawelpicode{code/fo_tm}
  \end{minipage}
  &
  \begin{minipage}{0.24\textwidth}
   \inputrawelpicode{code/ho_tm}
  \end{minipage}
  \end{tabular}\vspace{4pt}
  \caption{\Fo{} and \Ho language}\vspace{0.3em}
  \label{code:common-terms}
  \Description[code:common-terms]{code:common-terms}
\end{figure}

@@

\input{code/mem_types}

@@

\input{code/fo_subst}
\input{code/ho_subst}

@@

\subsection{Compilation}

@@

The auxiliary function \elpiIn{close-links} cannot be explained here,
che palle. \marginpar{fix}

@@

\subsection{Execution}

\subsection{Decompilation}
